Model Checking

### 2024-04-02 owen/Write Ups/04_model_checking.qmd
adding version for comparing winner effects over tinme

$$
\alpha_j \sim \mathcal{N}(\mu_2, 1)
$$
$$
\mu_2 \sim \mathcal{N}(0, 1)
$$
$$
\tau_2 \sim \mathcal{N}(0, 1)
$$

@@

$$
\beta_j \sim \mathcal{N}(\mu_1, 1)
$$
$$
\mu_1 \sim \mathcal{N}(0, 1)
$$
$$
\tau_1 \sim \mathcal{N}(0, 1)
$$

@@

__These predictors are not currently rescaled__.

@@

We fit this to games from all players in the 1700-1900 range. 
We looked only at standard rated bullet games (no tournaments, etc),
which was 675k games, ranging from 500 - 128k games per player.
Fitting this model takes many hours on a computing cluster.

@@

```{r 17-19-bullet-global, fig.path = "figures/"}
fit <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
                    "all_rated_bullet_model.RDS"))

users <- readRDS(here("results/", "Full_Fits",
                      "lichess1700-1900/",
                      "users_bullet.RDS"))

mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters 17-19 Bullet",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))

@@

- Here the mean for $\mu_2$ is -0.0483, corresponding to a 48% chance of 
winning as black against an equally ranked opponent. This parameter
is not clearly different from 0 here. Posterior SD is large.

@@

- The global estimate for $\mu_1$, the global winner effect, is 
around 0. Posterior SD is large here.

```{r winner-effects-17-19-bullet, fig.path = "figures/"}
#| fig-height: 9

@@

mcmc_hist(fit$draws("beta"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Player Winner Effects 17-19 Bullet")
```

@@

-The largest posterior mean here is 1.32. This is saying that if on
average you win 50% of your games, but have instead won 10 in a row, your
probability of winning the next would increase by ~16% over that baseline.
However most estimated effects are small, and most of these posterior
distributions contain 0.

```{r indiv-effects-17-19-bullet, fig.path = "figures/"}
#| fig-height: 9

names(players) <- paste0("alpha[", 1:length(users), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit$draws("alpha"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Individual Player Effects 17-19 Bullet")
```

@@

- Many are quite similar to the global parameter here.

@@

We fit this model also to all players rated between 2000 and 2200. 
With the same restrictions on the games this gives us 834k games.
We have between 1200 and 89k games for each of these players.

```{r global-effects-20-22-bullet, fig.path = "figures/"}

fit2 <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
                    "all_rated_bullet_model.RDS"))

mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters 20-22 Bullet",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

users_20_22 <- readRDS(here("results/", "Full_Fits",
                      "lichess2000-2200/",
                      "users_bullet.RDS"))

fit2$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))

@@

```{r winner-effects-20-22-bullet, fig.path = "figures/"}
#| fig-height: 9

@@

@@

mcmc_hist(fit2$draws("beta"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Player Winner Effects 20-22 Bullet")

@@

```{r indiv-effects-20-22-bullet, fig.path = "figures/"}
#| fig-height: 9

names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit2$draws("alpha"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Individual Player Effects, 20-22 Bullet")
```

@@

We look at all players rated between 2300 and 2500. This gives us 703k games,
with players playing 400-74k games.

```{r global-effects-23-25-bullet, fig.path = "figures/"}
fit3 <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
                    "all_rated_bullet_model.RDS"))

users_23_25 <- readRDS(here("results/", "Full_Fits",
                      "lichess2300-2500/",
                      "users_bullet.RDS"))

mcmc_hist(fit3$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters 23-25 Bullet",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit3$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
```

```{r winner-effects-23-25-bullet, fig.path = "figures/"}
#| fig-height: 9

@@

mcmc_hist(fit3$draws("beta"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Player Winner Effects 23-25 Bullet")

@@

```{r indiv-effects-23-25-bullet, fig.path = "figures/"}
#| fig-height: 9

names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit3$draws("alpha"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Individual Player Effects 23-25 Bullet")
```

@@

@@

Note that now there are 187k games here, considerably fewer than 
the number of bullet games played. This ranges from 35k games to 8 players
who play less than 25 blitz games.

@@

fit_blitz <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
                    "all_rated_blitz_model.RDS"))

mcmc_hist(fit_blitz$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters 17-19 Blitz",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit_blitz$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
```

@@

## winner effects
players <- users
names(players) <- paste0("beta[", 1:length(users), "]")

@@

mcmc_hist(fit_blitz$draws("beta"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Player Winner Effects, 17-19 Blitz")

@@

## indiv effects
names(players) <- paste0("alpha[", 1:length(users), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit_blitz$draws("alpha"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Individual Player Effects 17-19 Blitz")
```

@@

Here we have 250k games, ranging from 64k to 16 per player.

@@

fit2_blitz <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
                    "all_rated_blitz_model.RDS"))

mcmc_hist(fit2_blitz$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters 20-22 Blitz",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit2_blitz$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))

@@

## winner effects
players <- users_20_22
names(players) <- paste0("beta[", 1:length(users_20_22), "]")

@@

mcmc_hist(fit2_blitz$draws("beta"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Player Winner Effects 20-22 Blitz")

@@

## indiv effects
names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit2_blitz$draws("alpha"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Individual Player Effects 20-22 Blitz")

@@

Here we have only 82k total games, ranging from 15k to 16 games per player.

@@

fit3_blitz <- readRDS(here("results/", "Full_Fits", "lichess2300-2500/", 
                    "all_rated_blitz_model.RDS"))

mcmc_hist(fit3_blitz$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters 23-25 Blitz",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit3_blitz$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))

@@

## winner effects
players <- users_23_25
names(players) <- paste0("beta[", 1:length(users_23_25), "]")

@@

mcmc_hist(fit3_blitz$draws("beta"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Player Winner Effects, 23-25 Blitz")

@@

## indiv effects
names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit3_blitz$draws("alpha"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Individual Player Effects, 23-25 Blitz")

@@

fit_prev <- readRDS(here("results/", "Full_Fits", "lichess1700-1900/", 
                    "all_rated_bullet_model_prev.RDS"))

mcmc_hist(fit_prev$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters, 17-19 Bullet Previous",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit_prev$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
```

@@

## winner effects
players <- users
names(players) <- paste0("delta[", 1:length(users), "]")

@@

mcmc_hist(fit_prev$draws("delta"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Player Winner Effects, 17-19 Bullet Previous")
```

@@

## indiv effects
names(players) <- paste0("alpha[", 1:length(users), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit_prev$draws("alpha"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Individual Player Effects, 17-19 Bullet Previous")
```

@@

```{r 20-22-bullet-prev-global, fig.path="figures/"}
fit2_prev <- readRDS(here("results/", "Full_Fits", "lichess2000-2200/", 
                    "all_rated_bullet_model_prev.RDS"))

mcmc_hist(fit2_prev$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters 20-22 Bullet Previous",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit2_prev$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))
```

@@

## winner effects
players <- users_20_22
names(players) <- paste0("delta[", 1:length(users_20_22), "]")

@@

mcmc_hist(fit2_prev$draws("delta"),
          facet_args = list(labeller = player_labels, ncol = 5)) +
  labs(title = "Player Winner Effects 20-22 Bullet Previous")
```

>>>>

$$
\alpha_j \sim \mathcal{N}(0, \sigma_{\alpha}^2)
$$
$$
\sigma_{\alpha} \sim \mathcal{IG}(1, 1)
$$

@@

$$
\beta_j \sim \mathcal{N}(\mu_{\beta}, \sigma_{\beta}^2)
$$
$$
\mu_{\beta} \sim \mathcal{N}(0, \sigma^2)
$$
$$
\sigma_{\beta} \sim \mathcal{IG}(1, 1)
$$

and

$$
\sigma \sim \mathcal{IG}(1,1).
$$

@@

__These fixed effect predictors are not currently rescaled__.

@@

We fit this to games from all players in the 1700-1900 range. 
We looked only at standard rated bullet games (no tournaments, etc),
which was 590k games, ranging from 30 - 127k games per 36 players.
Fitting this model takes many hours on a computing cluster.

@@

```{r 17-19-bullet-global-update, fig.path = "figures/"}
fit <- readRDS(here("results/", "lichess1700-1900_mar/", 
                    "all_rated_bullet_model.RDS"))

users <- readRDS(here("results/",
                      "lichess1700-1900_mar/",
                      "users_bullet.RDS"))

fit$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))

mcmc_hist(fit$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
  labs(title = "Global Parameters 17-19 Bullet",
       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
       Gamma2 is ELO Difference, others SDs")

# mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
#   labs(title = "Global Parameters 17-19 Bullet",
#        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
#        Gamma2 is ELO Difference")
# 
# 
# fit$summary(variables = c("mu1", "mu2", "gamma1", "gamma2"))

@@

<!-- - Here the mean for $\mu_2$ is -0.0483, corresponding to a 48% chance of  -->
<!-- winning as black against an equally ranked opponent. This parameter -->
<!-- is not clearly different from 0 here. Posterior SD is large. -->

@@

- The global estimate for $\mu_{\beta}$, the global winner effect, is 
around 0. Posterior SD is large here.

```{r winner-effects-17-19-bullet-update, fig.path = "figures/"}
#| fig-height: 9

random_effect_post <- fit$draws() %>% as_draws_df() %>% 
  select(starts_with("beta[")) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]"))

@@

random_effect_post %>% 
  filter(param == 2) %>% 
  select(player_id, value) %>% 
  group_by(player_id) %>% 
  summarise(post_mean = mean(value)) %>% 
  arrange(-post_mean)

random_effect_post %>% 
  filter(param == 2) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 6) +
  labs(title = "Winner Effects", y = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

@@

-The largest posterior mean here is 0.4. This is saying that if on
average you win 50% of your games, but have instead won 10 in a row, your
probability of winning the next would increase by ~5% over that baseline.
However most estimated effects are small, and most of these posterior
distributions contain 0.

```{r indiv-effects-17-19-bullet-update, fig.path = "figures/"}
#| fig-height: 9

# names(players) <- paste0("alpha[", 1:length(users), "]")
# player_labels <- as_labeller(players)
# mcmc_hist(fit$draws("alpha"),
#           facet_args = list(labeller = player_labels, ncol = 5)) +
#   labs(title = "Individual Player Effects 17-19 Bullet")

random_effect_post %>% 
  filter(param == 1) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 6) +
  labs(title = "Individual Effects", y = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

@@

<!-- - Many are quite similar to the global parameter here. -->

@@

We fit this model also to all players rated between 2000 and 2200. 
With the same restrictions on the games this gives us 722k games.
We have between 150 and 88k games for each of these 39 players.

```{r global-effects-20-22-bullet-update, fig.path = "figures/"}

fit2 <- readRDS(here("results/", "lichess2000-2200_mar/", 
                    "all_rated_bullet_model.RDS"))

# mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
#   labs(title = "Global Parameters 20-22 Bullet",
#        subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
#        Gamma2 is ELO Difference")
mcmc_hist(fit2$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
  labs(title = "Global Parameters 20-22 Bullet",
       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
       Gamma2 is ELO Difference, others SDs")

users_20_22 <- readRDS(here("results/", 
                      "lichess2000-2200_mar/",
                      "users_bullet.RDS"))

fit2$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]", "tau[2]",
                           "sigma_1", "sigma_g1", "sigma_g2"))

@@

```{r winner-effects-20-22-bullet-update, fig.path = "figures/"}
#| fig-height: 9

@@

random_effect_post2 <- fit2$draws() %>% as_draws_df() %>% 
  select(starts_with("beta[")) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]"))

@@

random_effect_post2 %>% 
  filter(param == 2) %>% 
  select(player_id, value) %>% 
  group_by(player_id) %>% 
  summarise(post_mean = mean(value)) %>% 
  arrange(-post_mean)

random_effect_post2 %>% 
  filter(param == 2) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Winner Effects", y = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

```{r indiv-effects-20-22-bullet-update, fig.path = "figures/"}
#| fig-height: 9

random_effect_post2 %>% 
  filter(param == 1) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Individual Effects", y = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

# names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
# player_labels <- as_labeller(players)
# mcmc_hist(fit2$draws("alpha"),
#           facet_args = list(labeller = player_labels, ncol = 5)) +
#   labs(title = "Individual Player Effects, 20-22 Bullet")
```

@@

We look at all players rated between 2300 and 2500. This gives us 599k games,
with 38 players playing 17-74k games.

Size of the colour effect increases here when the ratings are larger,
as does the ELO difference effect.

```{r global-effects-23-25-bullet-update, fig.path = "figures/"}
fit3 <- readRDS(here("results/", "lichess2300-2500_mar/", 
                    "all_rated_bullet_model.RDS"))

users_23_25 <- readRDS(here("results/", 
                      "lichess2300-2500_mar/",
                      "users_bullet.RDS"))

mcmc_hist(fit3$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
  labs(title = "Global Parameters 23-25 Bullet",
       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
       Gamma2 is ELO Difference, others SDs")

fit3$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
```

```{r winner-effects-23-25-bullet-update, fig.path = "figures/"}
#| fig-height: 9

@@

random_effect_post3 <- fit3$draws() %>% as_draws_df() %>% 
  select(starts_with("beta[")) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]"))

random_effect_post3 %>% 
  filter(param == 2) %>% 
  select(player_id, value) %>% 
  group_by(player_id) %>% 
  summarise(post_mean = mean(value)) %>% 
  arrange(-post_mean)

random_effect_post3 %>% 
  filter(param == 2) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Winner Effects", y = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

```{r indiv-effects-23-25-bullet-update, fig.path = "figures/"}
#| fig-height: 9

random_effect_post3 %>% 
  filter(param == 1) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Individual Effects", y = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

# names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
# player_labels <- as_labeller(players)
# mcmc_hist(fit3$draws("alpha"),
#           facet_args = list(labeller = player_labels, ncol = 5)) +
#   labs(title = "Individual Player Effects 23-25 Bullet")
```

@@

NOT RERUN YET

@@

Note that now there are 109k games here, considerably fewer than 
the number of bullet games played. There are 8 players
who play less than 25 blitz games.

@@

fit_blitz <- readRDS(here("results/", "lichess1700-1900_mar/", 
                    "all_rated_blitz_model.RDS"))

users_blitz <- readRDS(here("results/",
                      "lichess1700-1900_feb/",
                      "users_blitz.RDS"))

mcmc_hist(fit_blitz$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
  labs(title = "Global Parameters 17-19 Blitz",
       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit_blitz$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
```

@@

# ## winner effects
# players <- users
# names(players) <- paste0("beta[", 1:length(users), "]")
# 
# player_labels <- as_labeller(players)
# 
# mcmc_hist(fit_blitz$draws("beta"),
#           facet_args = list(labeller = player_labels, ncol = 5)) +
#   labs(title = "Player Winner Effects, 17-19 Blitz")

random_effect_post <- fit_blitz$draws() %>% as_draws_df() %>% 
  select(starts_with("beta[")) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]"))

players <- users_blitz
names(players) <- paste0("beta[", 1:length(users_blitz), "]")

@@

random_effect_post %>% 
  filter(param == 2) %>% 
  select(player_id, value) %>% 
  group_by(player_id) %>% 
  summarise(post_mean = mean(value)) %>% 
  arrange(-post_mean)

random_effect_post %>% 
  filter(param == 2) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Winner Effects, 1700-1900 Blitz", y = "", x = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

## indiv effects
random_effect_post %>% 
  filter(param == 1) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Indiv Effects, 1700-1900 Blitz", y = "", x = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

# names(players) <- paste0("alpha[", 1:length(users), "]")
# player_labels <- as_labeller(players)
# mcmc_hist(fit_blitz$draws("alpha"),
#           facet_args = list(labeller = player_labels, ncol = 5)) +
#   labs(title = "Individual Player Effects 17-19 Blitz")
```

@@

Here we have 190k games.

@@

fit2_blitz <- readRDS(here("results/", "lichess2000-2200_feb/", 
                    "all_rated_blitz_model.RDS"))

users_blitz_2 <- readRDS(here("results/",
                      "lichess2000-2200_feb/",
                      "users_blitz.RDS"))

mcmc_hist(fit2_blitz$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
  labs(title = "Global Parameters 20-22 Blitz",
       subtitle = "mu_beta is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit2_blitz$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))

@@

random_effect_post2 <- fit2_blitz$draws() %>% as_draws_df() %>% 
  select(starts_with("beta[")) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]"))

players <- users_blitz_2
names(players) <- paste0("beta[", 1:length(users_blitz_2), "]")

@@

random_effect_post2 %>% 
  filter(param == 2) %>% 
  select(player_id, value) %>% 
  group_by(player_id) %>% 
  summarise(post_mean = mean(value)) %>% 
  arrange(-post_mean)

random_effect_post2 %>% 
  filter(param == 2) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Winner Effects, 2000-2200 Blitz", y = "", x = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

random_effect_post2 %>% 
  filter(param == 1) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Indiv Effects, 2000-2200 Blitz", y = "", x = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

Here we have only 62k total games.

@@

fit3_blitz <- readRDS(here("results/", "lichess2300-2500_feb/", 
                    "all_rated_blitz_model.RDS"))

users_blitz_3 <- readRDS(here("results/",
                      "lichess2300-2500_feb/",
                      "users_blitz.RDS"))

mcmc_hist(fit3_blitz$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
  labs(title = "Global Parameters 23-25 Blitz",
       subtitle = "mu_beta is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit3_blitz$summary(variables =  c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))

@@

## winner effects
players <- users_blitz_3
names(players) <- paste0("beta[", 1:length(users_blitz_3), "]")

@@

random_effect_post3 <- fit3_blitz$draws() %>% as_draws_df() %>% 
  select(starts_with("beta[")) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]"))

random_effect_post3 %>% 
  filter(param == 2) %>% 
  select(player_id, value) %>% 
  group_by(player_id) %>% 
  summarise(post_mean = mean(value)) %>% 
  arrange(-post_mean)

random_effect_post3 %>% 
  filter(param == 2) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Winner Effects, 2300-2500 Blitz", y = "", x = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

random_effect_post3 %>% 
  filter(param == 1) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 5) +
  labs(title = "Indiv Effects, 2300-2500 Blitz", y = "", x = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

fit_prev <- readRDS(here("results/", "lichess1700-1900_mar/", 
                    "all_rated_bullet_model_prev.RDS"))

## these will be the same users

mcmc_hist(fit_prev$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
  labs(title = "Global Parameters, 17-19 Bullet Previous",
       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit_prev$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))
```

The effects when only looking at the previous game are 
much smaller than previous 10, mostly very close to 0 here.

@@

## this needs to be updated now..

## winner effects
players <- users

names(players) <- paste0("beta[", 1:length(users), "]")

@@

random_effect_post4 <- fit_prev$draws() %>% as_draws_df() %>% 
  select(starts_with("beta[")) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]"))

random_effect_post4 %>% 
  filter(param == 2) %>% 
  select(player_id, value) %>% 
  group_by(player_id) %>% 
  summarise(post_mean = mean(value)) %>% 
  arrange(-post_mean)

random_effect_post4 %>% 
  filter(param == 2) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 6) +
  labs(title = "Winner Effects, 1700-1900 Bullet, Previous Game", y = "", x = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

The individual effects seem to be quite small here also.

@@

## indiv effects
random_effect_post4 %>% 
  filter(param == 1) %>% 
  mutate(id = as.numeric(id)) %>% 
  ggplot(aes(value)) +
  geom_histogram(fill = "#6497b1", colour = "black", linewidth = 0.2) +
  facet_wrap(~player_id, scales = "free", labeller = player_labels, 
             ncol = 6) +
  labs(title = "Indiv Effects, 1700-1900 Bullet Previous", y = "", x = "") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

@@

```{r 20-22-bullet-prev-global, fig.path="figures/"}
fit2_prev <- readRDS(here("results/", "lichess2000-2200_mar/", 
                    "all_rated_bullet_model_prev.RDS"))

mcmc_hist(fit2_prev$draws(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))) +
  labs(title = "Global Parameters, 20-22 Bullet Previous",
       subtitle = "mu_beta is Winner Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

fit2_prev$summary(variables = c("mu_beta", "gamma1", "gamma2", "tau[1]",
                          "tau[2]", "sigma_1", "sigma_g1", "sigma_g2"))

@@

## to be updated...
## winner effects
players <- users_20_22

names(players) <- paste0("beta[", 1:length(users), "]")

@@

random_effect_post5 <- fit2_prev$draws() %>% as_draws_df() %>% 
  select(starts_with("beta[")) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]"))

random_effect_post5 %>% 
  filter(param == 2) %>% 
  select(player_id, value) %>% 
  group_by(player_id) %>% 
  summarise(post_mean = mean(value)) %>% 
  arrange(-post_mean)
